
<!---
  ## 1.3. Auxiliary functions for `RSiena::SienaGOF()`

  Please see: `RSiena::SienaGOF()`.

I want to be able to have a dyadcensus/triadcensus in the GOF of Rsiena.
Thus, these functions have been added to `RsienaTwoStep`.
--->
  ```{r, echo=FALSE, eval=FALSE}
DyadCensus.sna <- function(i, data, sims, period, groupName, varName, levls=1:3){
  unloadNamespace("igraph") # to avoid package clashes
  require(network)
  require(sna)
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}
  # because else triad.census(x) will lead to an error
  tc <- sna::dyad.census(x)[levls]
  # names are transferred automatically
  tc
}

DyadCensus.sna <- cmpfun(DyadCensus.sna)


TriadCensus.sna <- function(i, data, sims, period, groupName, varName, levls=1:16){
  unloadNamespace("igraph") # to avoid package clashes
  require(network)
  require(sna)
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}

  # because else triad.census(x) will lead to an error
  tc <- sna::triad.census(x)[levls]
  # names are transferred automatically
  tc
}

TriadCensus.sna <- cmpfun(TriadCensus.sna)


Nacf.sna <- function(i, data, sims, period, groupName, varName, cov){
  unloadNamespace("igraph") # to avoid package clashes
  require(network)
  require(sna)
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}
  x <- as.sociomatrix.sna(x)
  snaM1 <- matrix(sna::nacf(x, cov, type = "moran", neighborhood.type = "out", demean = TRUE)[2]) #because I need dimension in order to set names
  snaM1
}

Nacf.sna <- cmpfun(Nacf.sna)
```


```{r, echo=FALSE, eval=FALSE}
#AAARGGHHH, It doesnt work. Okay, problem was that a vector of length 1 has no dimension, thus needed to convert to matrix.
# sims <- ans3$sims
# sims[[1]]
# sna::as.sociomatrix.sna(sims[[1]])
# #
# net <- networkExtraction(i=NULL, obsData=ans3$f, sims=ans3$sims, period=1, groupName="Data1", varName="mynet")
# #
# # sna::as.sociomatrix.sna(net)
# s502 == sna::as.sociomatrix.sna(net)
#
# x <- sna::as.sociomatrix.sna(net)
# sna::nacf(x, smoke, type = "moran", neighborhood.type = "out", demean = TRUE)[2]

# t <- (sna::nacf(x, smoke, type = "moran", neighborhood.type = "out", demean = TRUE)[2])
# rownames(t) <- "test"
#gives error

# t <- matrix(sna::nacf(x, smoke, type = "moran", neighborhood.type = "out", demean = TRUE)[2])
# rownames(t) <- "test"

# network.ts <- function(i, data, sims, period, groupName, varName){
#   unloadNamespace("igraph") # to avoid package clashes
#   require(network)
#   require(sna)
#   x <- networkExtraction(i, data, sims, period, groupName, varName)
#   x
# }
# network.ts <- cmpfun(network.ts)

```




```{r}
df_ts1m3 <- ts_nacf(sims=sims1m3, simtype="ministep", cov=mydata$cCovars$smoke)
names(df_ts1m3)[1] <- "MoranI"

```

```{r}
df <- rbind(dfrs, df_ts1m3)

p <- ggplot(df, aes(x=factor(type, levels=c("rsiena", "ministep")), y=MoranI, fill=factor(type, levels=c("rsiena", "ministep")))) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type")

p


```





```{r, echo=FALSE}
load("./data/processed/20231104ts_ansS.rda")
ts_ansS <- x
rm(x)
load("./data/processed/20231104ts_ansWC.rda")
ts_ansWC <- x
rm(x)
load("./data/processed/20231105ts_ansSC.rda")
ts_ansSC <- x
rm(x)
load("./data/processed/20231105ts_ansST.rda")
ts_ansST <- x
rm(x)
```




```{r}
ts_ansS <- ts_estim(net1 = s501, net2= s502, statistics = STATS, ccovar = DF, parallel = TRUE)

sims2m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), chain=FALSE)
fsave(sims2m3, file="sims2m3.Rdata")

#weak coordination
sims3m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), dist1=2, dist2=2, chain=FALSE)
fsave(sims3m3, file="sims3m3.Rdata")

#strict coordination
sims4m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), dist1=2, chain=FALSE)
fsave(sims4m3, file="sims4m3.Rdata")

#simstep
sims5m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,0,1), chain=FALSE)
fsave(sims5m3, file="sims5m3.Rdata")
```





```{r}
ts_ans <- ts_estim(ans = ans3, nite = 300, p2step=c(1,0,0), phase1 = TRUE, conv = 0.01, b=.7, parallel = TRUE)

ts_ans2 <- ts_estim(ans = ans3, nite = 300, p2step=c(0,1,0), dist1=2, phase1 = TRUE, conv = 0.01, b=.7, parallel = TRUE)

ts_phase1(mydata = mydata, myeff = myeff)
ts_ans <- ts_estim(mydata=mydata, myeff = myeff, nite = 300, p2step=c(1,0,0), phase1 = TRUE, conv = 0.01, b=.7)

ts_sim(mydata = mydata, myeff = myeff, preparedata = FALSE)
ts_phase1(mydata = mydata, myeff = myeff) #error in ts_targets
ts_estim(mydata = mydata, myeff = myeff, nite = 100, p2step=c(1,0,0), phase1 = TRUE) #error
ts_estim(net1 = s501, net2= s502, statistics = STATS, ccovar = DF, nite = 100, p2step=c(1,0,0), phase1 = TRUE)

```


`r colorize("These simulations took two days to complete on a simple HP Elitebook 855 G7", "red" )`

Prepare dataset for RsienaTwoStep Simulations.

```{r, eval=FALSE}
alcohol <- s50a[,1]
smoke <- s50s[,1]
df <- data.frame(alcohol=alcohol, smoke=smoke)
```


```{r, eval=FALSE, class.source="watch-out"}
#ministep
sims1m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=summary(ans3)$theta[2:9], p2step=c(1,0,0), chain=FALSE)

fsave(sims1m3, file="sims1m3.Rdata")
```



```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20230501sims1m3.Rdata")
sims1m3 <- x
rm(x)
```




## 5.3. Counting dyads

### 5.3.1. RSiena

```{r, message=FALSE}
gofi <- sienaGOF(ans3,
                 DyadCensus.sna,
                 verbose = TRUE,
                 join = TRUE,
                 cluster = my.cluster,
                 varName = "mynet")


```

```{r}
dfrs <- as.data.frame(gofi[[1]]$Simulations)
names(dfrs) <- c("Mut", "Asym", "Null")

df <- rbind(dfrs, dfrs, dfrs)
df$x <- rep(c("mut", "asym", "null"), each=nrow(dfrs))
df$y <- NA
df$y[df$x=="mut"] <- df$Mut[df$x=="mut"]
df$y[df$x=="asym"] <- df$Asym[df$x=="asym"]
df$y[df$x=="null"] <- df$Null[df$x=="null"]
df$type <- "rsiena"
dfrs_sienam3 <- df
rm(list=c("df", "dfrs"))
```

### 5.3.2. RsienaTwoStep

```{r}
df_ts1m3 <- ts_dyads(sims=sims1m3, simtype="ministep")
```

```{r}
df <- rbind(dfrs_sienam3, df_ts1m3)

p <- ggplot(df, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep"))) ) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```











## 5.4 Assessing segregation

### 5.4.1. Rsiena


```{r}
gofi <- sienaGOF(ans3,
                 Nacf.sna,
                 verbose = TRUE,
                 join = TRUE,
                 cluster = my.cluster,
                 varName = "mynet",
                 cov=mydata$cCovars$smoke)
```

```{r}
dfrs <- as.data.frame(gofi[[1]]$Simulations)
dfrs$type <- "rsiena"
names(dfrs)[1] <- "MoranI"
```

### 5.4.2. RsienaTwoStep



```{r}
df_ts1m3 <- ts_nacf(sims=sims1m3, simtype="ministep", cov=mydata$cCovars$smoke)
names(df_ts1m3)[1] <- "MoranI"

```

```{r}
df <- rbind(dfrs, df_ts1m3)

p <- ggplot(df, aes(x=factor(type, levels=c("rsiena", "ministep")), y=MoranI, fill=factor(type, levels=c("rsiena", "ministep")))) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type")

p


```
Okay, it seems RsienaTwoStep is programmed correctly.

# 6. Model 4: Model 3 with smoke segregation

## 6.1. Simulate networks via `RsienaTwoStep`

Let us now, simulate a model with segregation. Or more precisely, that pupils have a preference for a tie with someone with similar smoking behavior. We take as starting point the estimates of `ans3`.


```{r, eval=FALSE, class.source="watch-out"}
est <- summary(ans3)$theta[2:9]
est[5] <- .8 #count carefully :-(
est
```

Now simulate under the different twostep conditions.

Prepare dataset for RsienaTwoStep Simulations (again).

```{r, eval=FALSE}
alcohol <- s50a[,1]
smoke <- s50s[,1]
df <- data.frame(alcohol=alcohol, smoke=smoke)
```


```{r, eval=FALSE, class.source="watch-out"}
sims1m3b <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(1,0,0), chain=FALSE)

fsave(sims1m3b, file="sims1m3b.Rdata")

#simultaneity
sims2m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), chain=FALSE)
fsave(sims2m3, file="sims2m3.Rdata")

#weak coordination
sims3m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), dist1=2, dist2=2, chain=FALSE)
fsave(sims3m3, file="sims3m3.Rdata")

#strict coordination
sims4m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,1,0), dist1=2, chain=FALSE)
fsave(sims4m3, file="sims4m3.Rdata")

#simstep
sims5m3 <- ts_sims(nsims=100, parallel=TRUE, net=s501, ccovar=df, rate=summary(ans3)$theta[1], statistics=list(ts_degree, ts_recip, ts_transTrip, ts_cycle3, list(ts_simX, "smoke"), list(ts_altX, "alcohol"), list(ts_egoX, "alcohol"), list(ts_egoXaltX, "alcohol")), parameters=est, p2step=c(0,0,1), chain=FALSE)
fsave(sims5m3, file="sims5m3.Rdata")

```


```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20230501sims1m3.Rdata")
sims1m3 <- x
load("./data/processed/20230507sims1m3b.Rdata")
sims1m3b <- x
load("./data/processed/20230507sims2m3.Rdata")
sims2m3 <- x
load("./data/processed/20230507sims3m3.Rdata")
sims3m3 <- x
load("./data/processed/20230507sims4m3.Rdata")
sims4m3 <- x
load("./data/processed/20230507sims5m3.Rdata")
sims5m3 <- x
rm(x)
```



## 6.2 counting dyads RsienaTwoStep

```{r}
df_ts1m3 <- ts_dyads(sims=sims1m3, simtype="ministep-original")
df_ts1m3b <- ts_dyads(sims=sims1m3b, simtype="ministep-with segregation")
df_ts2m3 <- ts_dyads(sims=sims2m3, simtype="twostep-simultaneity")
df_ts3m3 <- ts_dyads(sims=sims3m3, simtype="twostep-weak coordination")
df_ts4m3 <- ts_dyads(sims=sims4m3, simtype= "twostep-strict coordination")
df_ts5m3 <- ts_dyads(sims=sims5m3, simtype="simstep")

```

### 6.2.1 Plotting all dyads

```{r}
df <- rbind(df_ts1m3, df_ts1m3b, df_ts2m3, df_ts3m3, df_ts4m3, df_ts5m3)

p <- ggplot(df, aes(x=x, y=y, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity","twostep-weak coordination", "twostep-strict coordination","simstep" ))) ) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```

### 6.2.1 Plotting dyads selection

```{r}
df_sel <- df[df$x=="asym" | df$x=="mut",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity","twostep-weak coordination", "twostep-strict coordination","simstep" ))) ) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```

## 6.3 counting triads RsienaTwoStep

```{r}
df_ts1m3 <- ts_triads(sims=sims1m3, simtype="ministep-original")
df_ts1m3b <- ts_triads(sims=sims1m3b, simtype="ministep-with segregation")
df_ts2m3 <- ts_triads(sims=sims2m3, simtype="twostep-simultaneity")
df_ts3m3 <- ts_triads(sims=sims3m3, simtype="twostep-weak coordination")
df_ts4m3 <- ts_triads(sims=sims4m3, simtype= "twostep-strict coordination")
df_ts5m3 <- ts_triads(sims=sims5m3, simtype="simstep")

```

### 6.3.1. Plotting all triad types

```{r}
df <- rbind(df_ts1m3, df_ts1m3b, df_ts2m3, df_ts3m3, df_ts4m3, df_ts5m3)

p <- ggplot(df, aes(x=x, y=y, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simstep")))) +
  geom_violin(position=position_dodge(1)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) +
  labs(x = "triad type", y = "triad count", fill="simulation type")

p
```

### 6.3.2 Plotting triad census selection I

```{r}
df_sel <- df[df$x=="030C" | df$x=="120C",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simstep")))) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "(cyclic) triad types", y = "triad count", fill="simulation type")

p
```
### 6.3.3 Plotting triad census selection II

```{r}
df_sel <- df[df$x=="210" | df$x=="300",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simstep")))) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "(cyclic) triad types", y = "triad count", fill="simulation type")

p
```


## 6.4 Assessing segregation RsienaTwoStep

```{r}
df_ts1m3 <- ts_nacf(sims=sims1m3, simtype="ministep-original", cov=mydata$cCovars$smoke)
names(df_ts1m3)[1] <- "MoranI"
df_ts1m3b <- ts_nacf(sims=sims1m3b, simtype="ministep-with segregation", cov=mydata$cCovars$smoke)
names(df_ts1m3b)[1] <- "MoranI"
df_ts2m3 <- ts_nacf(sims=sims2m3, simtype="twostep-simultaneity", cov=mydata$cCovars$smoke)
names(df_ts2m3)[1] <- "MoranI"
df_ts3m3 <- ts_nacf(sims=sims3m3, simtype="twostep-weak coordination", cov=mydata$cCovars$smoke)
names(df_ts3m3)[1] <- "MoranI"
df_ts4m3 <- ts_nacf(sims=sims4m3, simtype="twostep-strict coordination", cov=mydata$cCovars$smoke)
names(df_ts4m3)[1] <- "MoranI"
df_ts5m3 <- ts_nacf(sims=sims5m3, simtype="simstep", cov=mydata$cCovars$smoke)
names(df_ts5m3)[1] <- "MoranI"

```


```{r}
df <- rbind(df_ts1m3, df_ts1m3b, df_ts2m3, df_ts3m3, df_ts4m3, df_ts5m3)

p <- ggplot(df, aes(x=factor(type , levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity","twostep-weak coordination", "twostep-strict coordination","simstep")), y=MoranI, fill=factor(type, levels=c("ministep-original", "ministep-with segregation", "twostep-simultaneity","twostep-weak coordination", "twostep-strict coordination","simstep")))) +
  geom_violin(position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) +
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type")  +
  ylim(0, 1) +
  theme(axis.text.x=element_blank())

#  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type")
#+
#  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p



```


---


  # 7. Conclusion

  - The networks simulated by `RsienaTwoStep` assuming the ministep assumption lead to identical results as the simulated networks of `RSiena`. We can hence conclude that (at least for the used statistics) `RsienaTwoStep` is programmed correctly.
- For both model specifications, we observe that allowing for simultaneity does not really impact the observed dyads and triads in the network, although reciprocated ties seem a bit more likely. This is reassuring. A tentative conclusion would be that for relatively sparse and large networks it 'does not matter' to disregard simultaneous tie changes.
- Both forms of coordination (weak and strict) lead to more reciprocated ties.
- The dyad and triad census of the networks simulated assuming 'weak coordination' deviate the most from the other simulated networks. The assumption made under weak coordination is that also actors who are **`r colorize("not", "red")` connected at time t0** can initiate a coordinated tie-change. Hence, it is called weak coordination. However, these actors only evaluate twostep tie changes that lead to a connection at time t2. In other words, their choice set is reduced.
- Under strict coordination, it is assumed that only actors **connected at time t0** can start a coordination. Hence, it is called strict coordination. These actors evaluate all twostep tie changes (also those that would be the result of two independent ministeps). In our ABM we specified that there was only strict coordination (parameter `p2step=c(0,1,0`). This will probably mean that some actors made a lot more tie-changes than others. It may be worthwhile to mix twosteps with ministeps and simsteps (i.e., `p2step=c(0.33,0.33,0.34)`).
- Under the simstep assumption, we observe more asymmetric dyads. This, intuitively makes sense, because you cannot make a reciprocated tie all by yourself. More surprisingly is that also cyclic triads (030C and 120C) seem to occur relatively often. Whether or not this is solely the results of the abundance of asymetric dyads remains to be investigated.



---

