---
title: "Introduction_RsienaTwoStep"
output: rmarkdown::html_vignette
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Introduction_RsienaTwoStep}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)


```

```{r setup}
rm(list=ls())
library(RsienaTwoStep)
```

```{css, echo=FALSE}
.test {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
}

.test2 {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
  background-color: white;
  color: rgb(201, 76, 76);
}


h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
}

.button1 {
  background-color: grey; /* Red */ 
  border: 2px solid black;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  /* border-radius: 12px; */
  width: 100%;
}

.button1:hover {
  box-shadow: 0 12px 16px 0 rgba(0,0,0,0.24), 0 17px 50px 0 rgba(0,0,0,0.19);
}

.button1:active {
  border: 2px solid red;
}

pre {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 0px;
}

.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}

.large {
  max-height: 1100px;
  overflow-y: auto;
  overflow-x: auto;
  width: 1100px;
}


body{ /* Normal  */
    font-size: 14px;
}

h1 { /* Header 1 */
    font-size: 22px;
  font-weight: bold;
}

h2 { /* Header 2 */
    font-size: 20px;
  color: DarkBlue;
}

h3 { /* Header 3 */
    font-size: 18px;
  color: DarkBlue;
}

h4 { /* Header 4 */
    font-size: 16px;
  color: DarkBlue;
}

blockquote {
  padding: 10px 20px;
  margin: 0 0 20px;
  font-size: 14px;
  border-left: 6px solid #eee;
  background-color:lightcyan
}

caption {
  font-size: 14px;
  color: black;
  font-weight: bold
}

pre.numberSource a.sourceLine {
  left: -1em;
}

```


---  

# 1. Data sets

Let us have a look at the build-in data sets of `RsienaTwoStep`. 

## 1.1. ts_net1

### 1.1.1. Adjacency matrix

```{r }
ts_net1
```

### 1.1.2. Plot

```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=5, fig.cap="Figure 1.1.2. ts_net1"}
net1g <- igraph::graph_from_adjacency_matrix(ts_net1, mode="directed")
par(mar=c(.1, .1, .1, .1))
igraph::plot.igraph(net1g)
```

## 1.2. ts_net2

### 1.2.1. Adjacency matrix

```{r}
ts_net2
```

### 1.2.2. Plot

```{r, message=FALSE, warning=FALSE, fig.width=7, fig.height=5, fig.cap="Figure 1.2.2. ts_net2"}
net2g <- igraph::graph_from_adjacency_matrix(ts_net2, mode="directed")
par(mar=c(.1, .1, .1, .1))
igraph::plot.igraph(net2g)
```

---  

# 2. ABM ministep

## 2.1. Logic

1. sample ego  
2. construct possible alternative future networks based on all possible ministeps of ego    
3. calculate how sampled ego evaluates these possible networks  
4. Let the ego pick a network, that is, let agent decide on a tie-change  
5. GOTO 1 (STOPPING RULE: until you think we have made enough ministeps)


## 2.2. Possible networks after ministep  

Let us suppose we want to know what the possible networks are after all possible ministeps of one actor part of `ts_net2`. That is, let us assume that it is ego#2's turn to decide on tie-change. What are the possible networks? 

```{r}
options <- ts_alternatives_ministep(net=ts_net2, ego=2)
options
```

`ts_alternatives_ministep()` returns a list of all possible networks after all possible tie-changes available to ego#2 given network `ts_net2`. If you look closely you will see that `options[[2]]` equals the original network (i.e. ego#2 decided not to change any tie). 

## 2.3. Network statistics 

Which option will ego#2 choose? Naturally this will depend on which network characteristics (or statistics) ego#2 finds relevant. Let us suppose that ego#2 bases its decision solely on the number of ties it sends to others and the number of reciprocated ties it has with others. 

Let us count the number of ties ego#2 sends to alters. 

```{r}
ts_degree(net=options[[1]], ego=2)
```

And in the second (original) option: 

```{r}
ts_degree(net=options[[2]], ego=2)
```

In the package `RsienaTwoStep` there are functions for the following network statistics $s$: 

  - degree: `ts_degree()` 
  - reciprocity: `ts_recip()`  
  - outdegree activity: `ts_outAct()` 
  - indegree activity: `ts_inAct()` 
  - outdegree popularity: `ts_outPop()` 
  - indegree popularity: `ts_inPop()`  
  - transitivity: `ts_transTrip()` 
  - mediated transitivity: `ts_transMedTrip()`  
  - transitive reciprocated triplets: `ts_transRecTrip()` 
  - number of three-cycles: `ts_cycle3()`  
  
See @ripley2022manual (Chapter 12) for the mathematical formulation. Naturally, you are free to define your own network statistics. 

## 2.4. Evaluation function 

But what evaluation value does ego#2 attach to these network statistics and consequently to the network (in its vicinity) as a whole? Well these are the parameters, $\beta$, you will normally estimate with `RSiena::siena07()`. 
Let us suppose the importance for the statistic 'degree' is -1 and for the statistic 'reciprocity' is 2. 

So you could calculate the evaluation of the network saved in `options[[2]]` by hand: 

$$f^{eval}_{n}(\mathbf{s_{i}},\mathbf{\beta}) = \mathbf{s_i}^\mathsf{T}\mathbf{\beta}, $$
with $f^{eval}_{n}$ the evaluation function for agent $n$. $\mathbf{s_{i}}$ are the network statistics for network $i$ and $\mathbf{\beta}$ the corresponding parameters (or importance).

```{r}
1*-1 + 0*2
```

or with a little help of the network statistic functions:  

```{r}
ts_degree(net=options[[2]], ego=2)*-1 + ts_recip(net=options[[2]], ego=2)*2
```
Or you could use the `ts_eval()`. 

```{r}
eval <- ts_eval(net=options[[2]], ego=2, statistics=list(ts_degree, ts_recip), parameters=c(-1,2))
eval
```
Now, let us calculate the evaluation of all 5 possible networks: 
```{r}
eval <- sapply(options, FUN=ts_eval, ego=2, statistics=list(ts_degree, ts_recip), parameters=c(-1,2))
eval
```

## 2.5. Choice function 

So which option will ego#2 choose? Naturally this will be a stochastic process. But we see the last option has the highest evaluation. 
We use McFadden's choice function (for more information see [wiki](https://en.wikipedia.org/wiki/Discrete_choice)), that is let $P_{ni}$ be the probability that ego $n$ chooses network/alternative $i$. The choice function is then given by:  


$$P_{n,i} = \frac{exp(\mathbf{s_i}^\mathsf{T}\mathbf{\beta} )}{\Sigma_{j=1}^J exp(\mathbf{s_j}^\mathsf{T}\mathbf{\beta} )},$$  

with $s_i$ a vector of the value of each network statistics for network $i$ and $\beta$ is the vector of parameter values. Hence, $\mathbf{s_i}^\mathsf{T}\mathbf{\beta}$ is the value of the evaluation for network $i$.

Let us force ego#2 to make a decision. 

```{r, results='hold'}
choice <- sample(1:length(eval), size=1, prob=exp(eval)/sum(exp(eval)))
print("choice:") 
choice
print("network:")
options[[choice]]
```


If we repeat this process, that is...: 

1. sample agent  
2. construct possible alternative networks  
3. calculate how sampled agent evaluates the possible networks  
4. Let the agent pick a network, that is, let agent decide on a tie-change  
5. GO BACK TO 1 (STOPPING RULE: until you think we have made enough ministeps)

...we have an agent based model.   

## 2.6. Stopping rule  

But how many ministeps do we allow? Well, normally this is estimated by `siena07` by the `rate` parameter. If we do not make this rate parameter conditional on actor covariates or on network characteristics, the rate parameter can be interpreted as the average number of ministeps each actor in the network is allowed to make before time is up. Let us suppose the `rate` parameter is 2 `r rate<-2`. Thus in total the number of possible ministeps will be `nrow(ts_net2)*rate`: 10. For a more detailed - **and more correct** -  interpretation of the rate parameter in `siena07` see: [www.jochemtolsma.nl](www.jochemtolsma.nl/courses/complete-networks/socio6).

## 2.7. Example  

To demonstrate the network evolution:  

```{r}
#in startvalues the first value refers to the rate parameter, the subsequent values to the statistics of statistics

ts_sims(startvalues = c(2,-1,2), net1=ts_net2, statistics=list(ts_degree, ts_recip),nsims=1, p2step=c(1,0,0), chain = TRUE )
```

---  

# 3. ABM twostep

## 3.1. Logic

The general logic of the ABM that allows for twosteps is very similar to the ABM ministep model: 

1. sample **two** agents  
2. construct possible alternative networks  
3. calculate how the sampled agents evaluate the possible networks  
4. Let the agents together pick the subsequent network, that is, let agents decide on the twostep (the simultaneous two ministeps)  
5. GOTO 1 (STOPPING RULE: until you think we have made enough ministeps/twosteps)

## 3.2. Sample two agents

a. **Simultaneity**: agents are sampled randomly  
b. **Strict coordination**: only specific dyads are sampled (with a specific distance between them, based on either out-, in- or reciprocal ties)  
c. **Weak coordination**: agents are sampled randomly but only specific twosteps are regarded as 'coordinated' twosteps and, consequently only some of the possible alternative networks are included in the choice set for the dyad. 

## 3.3. Possible networks after twostep

If we want to allow for simultaneity, we simply let first agent1 make all possible ministeps and then conditional on these alternative networks let agent2 make all possible ministeps. Please note that the order in which we let agents make the ministeps is not important. We simple construct all the networks that could result from agent1 and agent2 make a simultaneous ministep.  

> <span style="color: red;">Exception:</span> 
> With weak coordination we will assess which possible alternative networks impact the evaluation function of both egos. Only those possible alternative networks are regarded as the result of coordination and included in the choice set. Thus, it is not necessarry for ego1 and ego2 to be connected at time1 but then they should at least be connected at time2 in such a way that they influence each others evaluation function.  
> The implementation in the current version of `RsieneTwoStep` is, however, a bit different. With weak coordination we simple assess the distance between ego1 and ego2 at time1 and time2. If at eithter time1 *or* time2 the distance is within the set threshold, we regard the twostep as a coordinated twostep.   

## 3.4. Network statistics  

We use the same network statistics as for the ABM ministep. But please note, that not all existing network statistics of `RSiena` make sense. See for example the figure below. Suppose the evaluation of the transitive triad for actor *i* depends on whether path *i* to *j* was the closing path (i.e., created after path *h* to *j*) or not. In a twostep, both ties may be created simultaneously and we cannot distinguish these two routes. 


```{r, echo=FALSE, fig.width=7, fig.height=5, results="hold", fig.cap="Figure 3.4. Twostep versus ministeps"}
triad <- matrix(rep(0,9), nrow=3, ncol=3)
triad[1,3] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
l <- igraph::layout.grid(triadg)
l[3,1] <- .5
l <- igraph::norm_coords(l, ymin=0, ymax=1, xmin=0, xmax=1)
triadg1 <- triadg

triad[3,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("grey", "red")
triadg2 <- triadg

triad[3,2] <- 0
triad[1,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("red", "grey")
triadg3 <- triadg

triad[3,2] <- 1
triad[1,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("red", "grey", "red")
triadg4 <- triadg

par(mfrow=c(3,3))
mat <- matrix(c(0,2,0,1,0,4,0,3,0), 3, byrow = TRUE)
layout(mat)
par(mar=c(.1, .1, .1, .1))
igraph::plot.igraph(triadg1, layout=l)
igraph::plot.igraph(triadg2, layout=l)
igraph::plot.igraph(triadg3, layout=l)
igraph::plot.igraph(triadg4, layout=l)
par(mfrow=c(1,1))
par(new = TRUE, ## add a new layer
            mar = c(0,0,0,0)) ## with no margins margins
## set up the plotting area for this layer
plot(1,1,xlim=c(0,1),ylim=c(0,1),type='n',xlab='',ylab='', frame.plot=FALSE, axes=FALSE) 
arrows(0.2,0.3,0.4,0.1, length=.15, lty=2) ## add arrow
arrows(0.2,0.7,0.4,0.9, length=.15, lty=2) ## add arrow
arrows(.6,0.9,.8,0.7, length=.15, lty=2) ## add arrow
arrows(0.6,0.1,0.8,0.3, length=.15, lty=2) ## add arrow
arrows(0.6,0.1,0.8,0.3, length=.15, lty=2) ## add arrow
arrows(0.3,0.5,0.7,0.5, length=.15, lty=3) ## add arrow


```
*note: Dashed arrows represent ministeps (long dash) or twostep (short dash); solid arrows represent initial ties (grey) or created ties (red).*


## 3.5. Evaluation function

We start by letting each involved agent evaluate all possible networks based on the individual evaluation functions. 
Thus agent1 gives an evaluation and agent2 gives an evaluation.  
Next we have to decide how to combine the separate evaluations of the two agents. For now, in `RsienaTwoStep`, we simply take the mean of the two separate evaluations as the final evaluation score. 

$$f^{eval}_{nm}(\mathbf{s_{i}},\mathbf{\beta}) = f^{eval}_{n}(\mathbf{s_{i}},\mathbf{\beta})  + f^{eval}_{m}(\mathbf{s_{i}},\mathbf{\beta})$$



## 3.6. Choice function

Here we follow the same logic as before. If we know the evaluation score of each network we simply apply a Mc Fadden's choice function. That is, the actors together 'decide' on the future network out of the possible alternative networks in the choice set, given the combined evaluation of these networks. Thus we see the dyad formed by agent1 and agent2 as the decision agent. Please note that agents (or rather the dyad) thus favor the network with the highest combined 'utility' score. This is not necessarily the network that would give one of the two agents the highest satisfaction.  


$$P_{nm,i} = \frac{exp(f^{eval}_{n}(\mathbf{s_{i}},\mathbf{\beta}) + f^{eval}_{m}(\mathbf{s_{i}},\mathbf{\beta})) }{\Sigma_{j=1}^J exp(f^{eval}_{n}(\mathbf{s_{j}},\mathbf{\beta}) + f^{eval}_{m}(\mathbf{s_{j}},\mathbf{\beta}))},$$  

## 3.7. Stopping rule

Once again the logic is exactly similar. However, we count a twostep as two ministeps. Thus, if each actor is allowed to make on average 8 ministeps, actors are allowed to make on average 4 twosteps. 

## 3.8. Example

```{r}
ts_sims(startvalues = c(2,-1,2), net1=ts_net2, statistics=list(ts_degree, ts_recip),nsims=1, p2step=c(0,1,0), chain = TRUE )
```

---  

# 4. ABM simstep

## 4.1. Logic  

A second way in which the ministep assumption can be relaxed is to allow the same actor to make two ministeps simultaneously. Suppose a triad formed by actors *i*, *j* and *h* and a tie between *h* and *j*. 


```{r, echo=FALSE, fig.width=7, fig.height=5, results="hold", fig.cap="Figure 4.1. Simstep versus ministeps"}
triad <- matrix(rep(0,9), nrow=3, ncol=3)
triad[3,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
l <- igraph::layout.grid(triadg)
l[3,1] <- .5
l <- igraph::norm_coords(l, ymin=0, ymax=1, xmin=0, xmax=1)
triadg1 <- triadg

triad[1,3] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("red", "grey")
triadg2 <- triadg

triad[1,3] <- 0
triad[1,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("red", "grey")
triadg3 <- triadg

triad[1,3] <- 1
triad[1,2] <- 1
triadg <- igraph::graph_from_adjacency_matrix(triad, mode="directed")
igraph::V(triadg)$label <- c("i", "j", "h")
igraph::V(triadg)$size <- 50
igraph::E(triadg)$color <- c("red", "red", "grey")
triadg4 <- triadg

par(mfrow=c(3,3))
mat <- matrix(c(0,2,0,1,0,4,0,3,0), 3, byrow = TRUE)
layout(mat)
par(mar=c(.1, .1, .1, .1))
igraph::plot.igraph(triadg1, layout=l)
igraph::plot.igraph(triadg2, layout=l)
igraph::plot.igraph(triadg3, layout=l)
igraph::plot.igraph(triadg4, layout=l)
par(mfrow=c(1,1))
par(new = TRUE, ## add a new layer
            mar = c(0,0,0,0)) ## with no margins margins
## set up the plotting area for this layer
plot(1,1,xlim=c(0,1),ylim=c(0,1),type='n',xlab='',ylab='', frame.plot=FALSE, axes=FALSE) 
arrows(0.2,0.3,0.4,0.1, length=.15, lty=2) ## add arrow
arrows(0.2,0.7,0.4,0.9, length=.15, lty=2) ## add arrow
arrows(.6,0.9,.8,0.7, length=.15, lty=2) ## add arrow
arrows(0.6,0.1,0.8,0.3, length=.15, lty=2) ## add arrow
arrows(0.3,0.5,0.7,0.5, length=.15, lty=3) ## add arrow
arrows(0.3,0.5,0.7,0.5, length=.15, lty=3) ## add arrow



```
*note: Dashed arrows represent ministeps (long dash) or simstep (short dash); solid arrows represent initial ties (grey) or created ties (red).*


Suppose, actors evaluate transitive triplet structures positively. Normally, under the ministep assumption, actor *i* should first make a tie to actor *h* (or *j*) and only when it is its turn again to make a tie change, make a tie to actor *j* (or *h*). Naturally, in larger networks a lot could have happened in the mean time. Given the network structure at the time, actor *i* is allowed to make the second ministep, actor *i* may not favor making the additional tie to actor *j* (or *h*) anymore.  
Also, actor *i* may not even want to create a tie to actor *h* or *j* if it is not already sure it can close the triad immediately afterwards (or even simultaneously). Naturally, in real life situations, it is not that strange to create multiple ties at (almost) the same time. See for example the paper on the impact of Kudos on running behavior [@FRANKEN2023151]. If you are on Strava (or on any other social media for that matter), it is very common to give multiple kudos (or likes) to different people in your network at - more or less - the same time.  

## 4.2. Possible networks after simstep  

Simply all networks that could arise after two sequential ministeps made by one actor (including the no change option). Please note that the choice set only contains unique network configurations.  

## 4.3. Network statistics  

We use the same network statistics as for the ABM ministep. But please note, that not all existing network statistics of `RSiena` make sense. See for example Figure 4.1. above. Suppose the evaluation of the transitive triad for actor *i* depends on whether path *i* to *j* was the closing path or the path *i* to *h*. In a simstep, both ties may be created simultaneously and we cannot distinguish these two routes.  

## 4.4. Evaluation function  

Identical to ABM ministep.  

## 4.5. Choice function.  

Identical to ABM ministep.

## 4.6. Stopping rule  

Once again the logic is exactly similar. However, we count a simstep as two ministeps. Thus if each actor is allowed to make on average 8 ministeps, actors are allowed to make on average 4 simsteps. 

## 4.7. Example

```{r}
ts_sims(startvalues = c(2,-2,2,2), net1=ts_net2, statistics=list(ts_degree, ts_recip, ts_outAct),nsims=1, p2step=c(0,0,1), chain = TRUE )
```


--- 


# 5. Network census

## 5.1. Simulate networks  

Let us simulate 100 times the outcome of a ABM twostep process and only save the final network

```{r}
nets <- ts_sims(startvalues = c(2,-2,2,2), net=ts_net2, statistics=list(ts_degree, ts_recip, ts_outAct),nsims=100, p2step=c(0,1,0), chain = FALSE, verbose = FALSE)
```

## 5.2. Dyad and triad configurations  

Now we want to count the dyad and triad configurations. 

```{r, warning=FALSE}
df_dyads <- ts_dyads(nets, forplot = FALSE, simtype="twostep: random")  
df_triads <- ts_triads(nets, forplot = FALSE, simtype="twostep: random") 
```

```{r, results="hold"}
df_dyads
df_triads
```

## 5.3. Violin plots  

If you want to make violin plots of these census, it is best to set the option `forplot` to `TRUE`. 

Just as an example: 

```{r}
df_dyads2 <- ts_dyads(nets, forplot = TRUE, simtype="twostep: random")  
```

```{r, fig.width=7, fig.height=5, fig.cap="Figure 5.3. Dyad census"}
library(ggplot2)
p <- ggplot(df_dyads2, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) + 
  xlab('dyad configuration') + ylab('count')
  
p

```

---  

# References

<!---note to self: two sequential ministeps is not the same as two simultaneous ministeps! RsienaTwoStep only implements simultaneity.--> 
