---
title: "MinistepToTwostep"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.test {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
}

.test2 {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
  background-color: white;
  color: rgb(201, 76, 76);
}


h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
}





.button1 {
  background-color: grey; /* Red */ 
  border: 2px solid black;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  /* border-radius: 12px; */
  width: 100%;
}

.button1:hover {
  box-shadow: 0 12px 16px 0 rgba(0,0,0,0.24), 0 17px 50px 0 rgba(0,0,0,0.19);
}

.button1:active {
  border: 2px solid red;
}

pre {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 0px;
}

.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}

.large {
  max-height: 1100px;
  overflow-y: auto;
  overflow-x: auto;
  width: 1100px;
}


body{ /* Normal  */
    font-size: 14px;
}

h1 { /* Header 1 */
    font-size: 20px;
  font-weight: bold;
}

h2 { /* Header 2 */
    font-size: 18px;
  color: DarkBlue;
}

h3 { /* Header 3 */
    font-size: 16px;
  color: DarkBlue;
}

blockquote {
  padding: 10px 20px;
  margin: 0 0 20px;
  font-size: 14px;
  border-left: 6px solid #eee;
  background-color:lightcyan
}

caption {
  font-size: 14px;
  color: black;
  font-weight: bold
}

pre.numberSource a.sourceLine {
  left: -1em;
}

```

# 1. Getting started


```{r}
rm(list=ls())
```


## 1.1. Basic functions  


```{r}
fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}

```

## 1.2. Packages

```{r}
packages = c("RsienaTwoStep", "doParallel", "ggplot2")

fpackage.check(packages)
```

---  

# 2. Setting up cluster

```{r}
#stopCluster(my.cluster) 
n.cores <- parallel::detectCores() - 1  #save one core for other work
# create the cluster
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK")
# register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)
```

---  

# 3. Running simulations on toy data

## 3.1. Model 1: Degree and reciprocity 

Let us assume people really don't like to have a non-reciprocal tie but do like reciprocal ties a lot. (I only want to help you if you help me!)

Thus if you start with a network without many (reciprocal) ties it would be very difficult to get more reciprocal ties in the normal model. However, with simultaneity this should be possible. 

### 3.1.1. Simulate networks 

```{r, eval=FALSE}
sims1 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip), parameters=c(-1,2), p2step=0, chain=FALSE) #ministep

sims2 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip), parameters=c(-1,2), p2step=1, chain=FALSE) #twostep: simultaneity

sims3 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip), parameters=c(-1,2), p2step=1, dist1=2, dist2=2, chain=FALSE) #twostep: weak coordination

sims4 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip), parameters=c(-1,2), p2step=1, dist1=2,  chain=FALSE) #twostep: strict coordination
```

### 3.1.2. Counting dyads

```{r, eval=FALSE}
df1 <- ts_dyads(sims=sims1, simtype="ministep")
df2 <- ts_dyads(sims=sims2, simtype="twostep: simultaneity")  
df3 <- ts_dyads(sims=sims3, simtype="twostep: weak coordination")  
df4 <- ts_dyads(sims=sims4, simtype="twostep: strict coordination")  

df <- rbind(df1, df2, df3, df4)
```

```{r, eval=FALSE, echo=FALSE}
fsave(df, "dfm1")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221216dfm1")
df <- x
rm(x)
```

### 3.1.3. Plot results of dyadcensus 

```{r}
p <- ggplot(df, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 
  
p

```

### 3.1.4. Conclusion

Yes, if you allow simultaneity, we will observe more reciprocal ties after same number of possible tie-changes. 

Perhaps the model with ministeps only would reach the same target values (for dyads) but probably by using more possible tie-changes. This would mean there will be more noise in the network chains, this can/will have results for other network statistics. 

Also note, that if you make the network larger, simultaneity will less likely to be different from two consecutive ministeps, this is because the two agents do not influence each others evaluation functions. 

---  


## 3.2. Model 2: Degree, reciprocity, transTrip 

Let us assume people really don't like to have a non-reciprocal tie but do like reciprocal ties a lot and do favor transitivity. 

Is it unlikely that a transitive triad will emerge if we have coordination, because for one of the actors will not evaluate the transitive triad positively because in wrong position. Or, because reciprocal ties are more likely when actors coordinate, we will also observe more transitive triads. 


### 3.2.1. Simulate networks 

```{r, eval=FALSE}
sims1 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=5, statistics=list(ts_degree, ts_recip, ts_transTrip), parameters=c(-2,2, 1), p2step=0, chain=FALSE) #ministep

sims2 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=5, statistics=list(ts_degree, ts_recip,ts_transTrip), parameters=c(-2,2, 1), p2step=1, chain=FALSE) #twostep: simultaneity

sims3 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=5, statistics=list(ts_degree, ts_recip,ts_transTrip), parameters=c(-2,2, 1), p2step=1, dist1=2, dist2=2, chain=FALSE) #twostep: weak coordination

sims4 <- ts_sims(nsims=1000, parallel=TRUE, net=net1, rate=5, statistics=list(ts_degree, ts_recip, ts_transTrip), parameters=c(-2,2,1), p2step=1, dist1=2,  chain=FALSE) #twostep: strict coordination

```

### 3.2.2. Counting dyads

```{r, eval=FALSE}
df1 <- ts_dyads(sims=sims1, simtype="ministep")
df2 <- ts_dyads(sims=sims2, simtype="twostep: simultaneity")  
df3 <- ts_dyads(sims=sims3, simtype="twostep: weak coordination")  
df4 <- ts_dyads(sims=sims4, simtype="twostep: strict coordination")  

df <- rbind(df1, df2, df3, df4)
```

```{r, eval=FALSE, echo=FALSE}
fsave(df, "dfm2")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221216dfm2")
df <- x
rm(x)
```

### 3.2.3. Plot results of the dyadcensus 

```{r}
p <- ggplot(df, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 
  
p


```

### 3.2.4. Counting triads

```{r, eval=FALSE}
df1 <- ts_triads(sims=sims1, simtype="ministep")
df2 <- ts_triads(sims=sims2, simtype="twostep: simultaneity")  
df3 <- ts_triads(sims=sims3, simtype="twostep: weak coordination")  
df4 <- ts_triads(sims=sims4, simtype="twostep: strict coordination")  

df <- rbind(df1, df2, df3, df4)
```

```{r, eval=FALSE, echo=FALSE}
fsave(df, "dfm2_t")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221216dfm2_t")
df <- x
rm(x)
```


### 3.2.5. Plot results of the triadcensus 

```{r, results='hold'}
# 030T *1
# 120U *1
# 120D *2
# 120C *2
# 210 * 4
# 300 * 6


dftest <- df[df$x=="030T" | df$x=="120U", ]
p1 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 

p1

dftest <- df[df$x=="120D" | df$x=="120C" , ]
p2 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 

p2

dftest <- df[df$x=="210" | df$x=="300" , ]
p3 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 

p3

```

## 3.3. Model 3: Degree, reciprocity, transTrip, transMedTrip

Let us assume people really don't like to have a non-reciprocal tie but do like reciprocal ties a lot and do favor transitivity. 

It is now likely that a transitive triad will emerge if we have coordination, because one of the actors evaluate the transTrip and the other may positively evaluate transMedTrip positively. 


### 3.3.1. Simulate networks

```{r, eval=FALSE}
nsims <- 500

sims1 <- ts_sims(nsims=nsims, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip, ts_transTrip, ts_transMedTrip), parameters=c(-2,1,2,2), p2step=0, chain=FALSE) #twostep: ministep

sims2 <- ts_sims(nsims=nsims, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip, ts_transTrip, ts_transMedTrip), parameters=c(-2,1,2,2), p2step=1, chain=FALSE) #twostep: random

sims3 <- ts_sims(nsims=nsims, parallel=TRUE, net=net1, rate=10, statistics=list(ts_degree, ts_recip, ts_transTrip, ts_transMedTrip), parameters=c(-2,1,2,2), p2step=1, dist1=2, dist2=2, chain=FALSE) #twostep: weak

sims4 <- ts_sims(nsims=nsims, parallel=TRUE, net=net1, rate=5, statistics=list(ts_degree, ts_recip, ts_transTrip, ts_transMedTrip), parameters=c(-2,1,2,2), p2step=1, dist1=2,  chain=FALSE) #twostep: strict coordination
```

### 3.3.2. Counting dyads

```{r, eval=FALSE}
df1 <- ts_dyads(sims=sims1, simtype="ministep")
df2 <- ts_dyads(sims=sims2, simtype="twostep: simultaneity")  
df3 <- ts_dyads(sims=sims3, simtype="twostep: weak coordination")  
df4 <- ts_dyads(sims=sims4, simtype="twostep: strict coordination")  

df <- rbind(df1, df2, df3, df4)
```

```{r, eval=FALSE, echo=FALSE}
fsave(df, "dfm3")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221216dfm3")
df <- x
rm(x)
```

### 3.3.3. Plot results of the dyadcensus 

```{r}
p <- ggplot(df, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 
  
p


```


### 3.3.4. Counting triads

```{r, eval=FALSE}
df1 <- ts_triads(sims=sims1, simtype="ministep")
df2 <- ts_triads(sims=sims2, simtype="twostep: simultaneity")  
df3 <- ts_triads(sims=sims3, simtype="twostep: weak coordination")  
df4 <- ts_triads(sims=sims4, simtype="twostep: strict coordination")  

df <- rbind(df1, df2, df3, df4)
```

```{r, eval=FALSE, echo=FALSE}
fsave(df, "dfm3_t")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221216dfm3_t")
df <- x
rm(x)
```


### 3.3.5. Plot results of the triadcensus 

```{r, results='hold'}
# Basic violin plot
# 030T *1
# 120U *1
# 120D *2
# 120C *2
# 210 * 4
# 300 * 6
#dftest <- df[df$x=="030T" | df$x=="120D" | df$x=="120U" | df$x=="120C" | df$x=="210" | df$x=="300", ]

dftest <- df[df$x=="030T" | df$x=="120U", ]
p1 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 
p1 

dftest <- df[df$x=="120D" | df$x=="120C" , ]
p2 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 
p2

dftest <- df[df$x=="210" | df$x=="300" , ]
p3 <- ggplot(dftest, aes(x=x, y=y, fill=type)) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) 

p3

```

--- 
