---
title: "RSienaWorkFlow"
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.test {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
}

.test2 {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
  background-color: white;
  color: rgb(201, 76, 76);
}


h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
}





.button1 {
  background-color: grey; /* Red */ 
  border: 2px solid black;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  /* border-radius: 12px; */
  width: 100%;
}

.button1:hover {
  box-shadow: 0 12px 16px 0 rgba(0,0,0,0.24), 0 17px 50px 0 rgba(0,0,0,0.19);
}

.button1:active {
  border: 2px solid red;
}

pre {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 0px;
}

.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}

.large {
  max-height: 1100px;
  overflow-y: auto;
  overflow-x: auto;
  width: 1100px;
}


body{ /* Normal  */
    font-size: 14px;
}

h1 { /* Header 1 */
    font-size: 20px;
  font-weight: bold;
}

h2 { /* Header 2 */
    font-size: 18px;
  color: DarkBlue;
}

h3 { /* Header 3 */
    font-size: 16px;
  color: DarkBlue;
}

blockquote {
  padding: 10px 20px;
  margin: 0 0 20px;
  font-size: 14px;
  border-left: 6px solid #eee;
  background-color:lightcyan
}

caption {
  font-size: 14px;
  color: black;
  font-weight: bold
}

pre.numberSource a.sourceLine {
  left: -1em;
}

```

# 1. Getting started


```{r}
rm(list=ls())
```


## 1.1. Basic functions  


```{r}
fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 3, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "600px")
}

```

## 1.2. Packages

```{r, results='hide', message=FALSE}
packages = c("RsienaTwoStep", "RSiena", "doParallel", "compiler", "ggplot2", "tidyverse", "kableExtra")

fpackage.check(packages)
```


---  

# 2. Setting up cluster

```{r, results='hide'}
no_cores <- detectCores()
mycl <- makeCluster(rep("localhost", no_cores))
clusterEvalQ(mycl, library(RsienaTwoStep))
clusterEvalQ(mycl, library("network"))
clusterEvalQ(mycl, library("RSiena"))
clusterEvalQ(mycl, library("sna"))
registerDoParallel(mycl)
#stopCluster(cl = mycl)
```


---  

# 3. Running `Siena07()`

## 3.1. Prepare the dataset
```{r}
friend.data.w1 <- s501
friend.data.w2 <- s502
drink <- s50a
smoke <- s50s

friendship <- sienaDependent(array(c(friend.data.w1, friend.data.w2), dim = c(50, 50,
    2)))  
alcohol <- sienaDependent(drink[,1:2], type = "behavior")
smoke <- coCovar(smoke[, 1])

mydata <- sienaDataCreate(friendship, smoke, alcohol)
#print01Report(mydata)
```

## 3.2. Set up up the algorithm. 

Set conditional to `FALSE`, this way we estimate the rate parameter and I will be able to retrieve the rate parameter estimate in theta. 
Also set 'findiff' to `TRUE`. In `RsienaTwoStep` the estimates of derivatives (phase1 and phase3) are estimated using finite differences. 

```{r, eval=FALSE, message=FALSE} 
myalgorithm <- sienaAlgorithmCreate(cond = FALSE, findiff = TRUE, projname=NULL)
```
## 3.3. Define the model
```{r}
myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, cycle3, transTrip)
myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
myeff <- includeEffects(myeff, absDiffX, interaction1 = "alcohol")
myeff <- includeEffects(myeff, name = "alcohol", avAlt, interaction1 = "friendship")
myeff <- includeEffects(myeff, name = "alcohol", effFrom, interaction1 = "smoke")
```

## 3.4. Estimate the model. 
```{r, eval=FALSE}
ansb1 <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, returnDeps = TRUE)
#fsave(ansb1, file="ansb1.Rdata")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20231221ansb1.Rdata")
ansb1 <- x
rm(x)
```

let's have a look
```{r}
ansb1
```
<br> 

# 4. Estimate via `RsienaTwoStep`

## 4.1. Prepare the dataset

```{r}
DF <- data.frame(alcohol.1 = s50a[, 1], alcohol.2 = s50a[, 2], smoke = s50s[, 1])
```


## 4.2. Define the model. 

Include all names of the statistics in a list. The aim is to use the shortname of these effects as listed in the RSiena model with a prefix "ts_" added to them. If the statistic requires a covariate use a list with the first element the name of the statistic and the second element the name of the covariate. This should be the same name as used in your dataset. 

```{r}
STATS <- list(ts_degree, 
              ts_recip, 
              ts_transTrip,
              ts_cycle3,
              list(ts_simX, "smoke"), 
              list(ts_absDiffX, "alcohol"), 
              ts_linear,
              ts_quad,
              list(ts_avAlt, "friendship"), 
              list(ts_effFrom, "smoke"))
```

### 4.2.1. Check statistics 

As a brief intermezzo, check if the statistics are programmed correctly by comparing the target values. 

```{r}
t1 <- ts_targets(ansb1) #target values calculated by `siena07()`
t2 <- ts_targets(mydata=mydata, myeff=myeff) #target values calculated by RsienaTwoStep based on RSiena objects. 
t3 <- ts_targets(net1 = s501, net2 = s502, statistics = STATS, ccovar = DF) #target values calculated by RsienaTwoStep based on RsienaTwoStep objects.
df <- data.frame(Siena_original = t1, ts_siena_objects = t2, ts_twostep_object = t3 )
rownames(df) <- names(t3)
fshowdf(df)
```
<br> 

Everything is fine. 

## 4.3. Estimate model 

### 4.3.1. Estimate parameters. 

We could estimate the parameters and the SE separately. This means we need to set `phase3` to `FALSE`. 

```{r, eval=FALSE}
ts_ansb1 <- ts_estim(net1 = s501, net2= s502, statistics = STATS, ccovar = DF, parallel = TRUE, phase3 = FALSE)
```

```{r, eval=FALSE, echo=FALSE}
fsave(ts_ansb1, file="ts_ansb1.Rdata")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20240107ts_ansb1.Rdata")
ts_ansb1 <- x
rm(x)
```

### 4.3.2. check convergence visually  

let's have a look if parameters converged. 

```{r}
df <- data.frame(ts_ansb1)
vars <- colnames(df)
df$ite <- 1:nrow(ts_ansb1)
#convert data from wide to long format
df <- df %>% pivot_longer(cols= all_of(vars),
                          names_to='statistic',
                          values_to='estimate')

ggplot(df, aes(x=ite, y=estimate)) + 
  geom_line(aes(color=statistic))
```
<br> 

### Estimates

```{r}
df_est <- data.frame(colMeans(ts_ansb1[750:1000,]))
df_est$RSiena <- ansb1$theta
colnames(df_est) <- c("RsienaTwoStep", "RSiena")
fshowdf(df_est)

```

<br> 

### 4.3.3. Estimate SE, tstats and tconv.max

```{r}
ESTIM <- df_est$RsienaTwoStep
```

```{r, eval=FALSE}
ansb1_phase3 <- ts_phase3(startvalues = ESTIM, net1 = s501, net2= s502, statistics = STATS, ccovar = DF, itef3 = 1000, parallel = TRUE, returnDeps = TRUE, verbose = TRUE)
```

```{r echo=FALSE, eval=FALSE}
fsave(ansb1_phase3, "ansb1_phase3.rda")
```

```{r, echo=FALSE}
load("./data/processed/20231025ansb1_phase3.rda")
ans1c_phase3 <- x
rm(x)
```

## 4.4 final results  

Let us have a look at the final results. 

```{r}
SE <- sqrt(diag(ansb1_phase3$covtheta))
tstat <- ansb1_phase3$tstat
tconv.max <- ansb1_phase3$tconv.max
df <- data.frame(estim = ESTIM, SE = SE, "tratio" = tstat)

knitr::kable(df, digits = 3, "html", escape=FALSE, col.names = c("Estimate", "Standard <br> Error", "Convergence <br> t-ratio")) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "500px") %>%
        kableExtra::footnote(general = paste("tconv.max:", round(tconv.max, 3), sep=" "))

```

<br> 
My conclusion is that with RSienatwostep we can properly estimate a (very simple) network co-evolution model using the common ministep assumption. 
Good job! 

# 5. Compare the estimates of the different twostep models


## 5.2. Estimate twostep models
### 5.2.1. phase2

We will use the ans of 'RSiena07()` as our input. We only estimate phase2. 
The goal is not so much to come to a perfect estimate but to check if the estimates are within the CI of the original estimates of RSiena. Only if this would not be the case, there may be a need to estimate the model by using different assumptions. 

```{r, eval=FALSE}
### simultaneity
ts_ansb1S <- ts_estim(ans = ans3, nite = 500, conv = 0.01, parallel = TRUE, phase3 = FALSE, p2step = c(0,1,0))
fsave(ts_ansb1S, "ts_ansb1S.rda")

### weak coordination
ts_ansb1WC <- ts_estim(ans = ans3, nite = 500, conv = 0.01, parallel = TRUE, phase3 = FALSE, p2step = c(0,1,0), dist1 = 2, dist2 = 2, modet1 = "degree", modet2 = "degree")
fsave(ts_ansb1WC, "ts_ansb1WC.rda")

### strict coordination
ts_ansb1SC <- ts_estim(ans = ans3, nite = 500, conv = 0.01, parallel = TRUE, phase3 = FALSE, p2step = c(0,1,0), dist1 = 2, modet1 = "degree" )
fsave(ts_ansb1SC, "ts_ansb1SC.rda")

### simstep
ts_ansb1ST <- ts_estim(ans = ans3, nite = 500, conv = 0.01, parallel = TRUE, phase3 = FALSE, p2step = c(0,0,1))
fsave(ts_ansb1ST, "ts_ansb1ST.rda")
```


```{r, echo=FALSE}
load("./data/processed/20231104ts_ansS.rda")
ts_ansS <- x
rm(x)
load("./data/processed/20231104ts_ansWC.rda")
ts_ansWC <- x
rm(x)
load("./data/processed/20231105ts_ansSC.rda")
ts_ansSC <- x
rm(x)
load("./data/processed/20231105ts_ansST.rda")
ts_ansST <- x
rm(x)
```

## 5.3. check convergence visually  {.tabset .tabset-fade} 


### 5.3.1. simultaneity
```{r, warning=FALSE, message=FALSE}
ts_ans <- ts_ansS
df <- data.frame(ts_ans)
vars <- colnames(df)
df$ite <- 1:nrow(ts_ans)
#convert data from wide to long format
df <- df %>% pivot_longer(cols= vars,
                          names_to='statistic',
                          values_to='estimate')

ggplot(df, aes(x=ite, y=estimate)) + 
  geom_line(aes(color=statistic))
```
<br> 
### 5.3.2. weak coordination
```{r}
ts_ans <- ts_ansWC
df <- data.frame(ts_ans)
vars <- colnames(df)
df$ite <- 1:nrow(ts_ans)
#convert data from wide to long format
df <- df %>% pivot_longer(cols= vars,
                          names_to='statistic',
                          values_to='estimate')

ggplot(df, aes(x=ite, y=estimate)) + 
  geom_line(aes(color=statistic))
```
<br> 
### 5.3.3. strict coordination
```{r}
ts_ans <- ts_ansSC
df <- data.frame(ts_ans)
vars <- colnames(df)
df$ite <- 1:nrow(ts_ans)
#convert data from wide to long format
df <- df %>% pivot_longer(cols= vars,
                          names_to='statistic',
                          values_to='estimate')

ggplot(df, aes(x=ite, y=estimate)) + 
  geom_line(aes(color=statistic))
```
<br> 
### 5.3.4. simstep
```{r}
ts_ans <- ts_ansST
df <- data.frame(ts_ans)
vars <- colnames(df)
df$ite <- 1:nrow(ts_ans)
#convert data from wide to long format
df <- df %>% pivot_longer(cols= vars,
                          names_to='statistic',
                          values_to='estimate')

ggplot(df, aes(x=ite, y=estimate)) + 
  geom_line(aes(color=statistic))
```
<br> 

## 5.4. Compare estimates 

```{r}
#CI of RSiena
down <- ans3$theta - 1.96 * sqrt(diag(ans3$covtheta))
up <- ans3$theta + 1.96 * sqrt(diag(ans3$covtheta))

#our statistics
stats <-  factor(vars, levels = vars)
#final estimates 
b_s <- ts_ansS[nrow(ts_ansS),]
b_sc <- ts_ansSC[nrow(ts_ansSC),]
b_wc <- ts_ansWC[nrow(ts_ansWC),]
b_st <- ts_ansST[nrow(ts_ansST),]

#put everything in a dataframe
my.dt <- data.frame(statistics = stats, estimate=ans3$theta, down=down, up=up, b_s=b_s, b_sc=b_sc, b_wc=b_wc, b_st=b_st)

#use different layers to plot the separate estimates. and include a manual legend. 
ggplot(my.dt, aes(x=statistics, y=estimate)) +         
      geom_point(size = 3, aes(color="ministep")) +
      geom_errorbar(aes(ymin = down, ymax = up, color="ministep")) + 
      geom_point(size = 3, alpha = 0.5, aes(x=stats, y=b_s, color="simultaneity" )) + 
      geom_point(size = 3, alpha = 0.5, aes(x=stats, y=b_sc , color="strict coordination")) + 
  geom_point(size = 3, alpha = 0.5, aes(x=stats, y=b_wc, color="weak coordination")) + 
  geom_point(size = 3, alpha = 0.5, aes(x=stats, y=b_st, color="simstep")) + 
  scale_color_manual(name= "Estimation type", 
                     breaks = c("ministep", "simultaneity", "strict coordination", "weak coordination", "simstep"),
                     values = c("ministep"= "black", "simultaneity" = "red", "strict coordination" = "blue", "weak coordination" = "green", "simstep" = "orange")) + 
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(n.breaks=10)

```
<br> 
This figure already tells us that our twostep model more or less leads to similar estimates. 
There are some notably exceptions, however. 

Weak coordination: much smaller degree estimate, much larger reciprocity estimate and larger transTrip estimate. This is not that strange since that under weak coordination the two alters have to be connected either at the beginning of the twostep or after the twostep.  


## 5.5. phase3 for all twostep models

Please note, this can take a long time (more than a week), even with the relatively low `itef3` default value of 100 in `ts_phase3()`. 

```{r, eval=FALSE}
b_s <- ts_ansS[nrow(ts_ansS),]
b_sc <- ts_ansSC[nrow(ts_ansSC),]
b_wc <- ts_ansWC[nrow(ts_ansWC),]
b_st <- ts_ansST[nrow(ts_ansST),]

### simultaneity
ts_ansSp3 <- ts_phase3(mydata = mydata, myeff = myeff, startvalues =  b_s, parallel = TRUE, returnDeps = TRUE, p2step = c(0,1,0))
fsave(ts_ansSp3, "ts_ansSp3.rda")

### weak coordination
ts_ansWCp3 <- ts_phase3(mydata = mydata, myeff = myeff, startvalues =  b_wc, parallel = TRUE, returnDeps = TRUE, p2step = c(0,1,0), dist1 = 2, dist2 = 2, modet1 = "degree", modet2 = "degree")
fsave(ts_ansWCp3, "ts_ansWCp3.rda")

### strict coordination
ts_ansSCp3 <- ts_phase3(mydata = mydata, myeff = myeff, startvalues =  b_sc, parallel = TRUE, returnDeps = TRUE, p2step = c(0,1,0), dist1 = 2, modet1 = "degree")
fsave(ts_ansSCp3, "ts_ansSCp3.rda")

### simstep
ts_ansSTp3 <- ts_phase3(mydata = mydata, myeff = myeff, startvalues =  b_st, parallel = TRUE, returnDeps = TRUE, p2step = c(0,0,1))
fsave(ts_ansSTp3, "ts_ansSTp3.rda")
```

### 5.5.1 final results! 

We conclude that all five different theories of interaction lead to similar conclusions with respect to significance of included statistics.  

```{r, echo=FALSE}
load("./data/processed/20231108ts_ansSp3.rda")
ts_ansSp3 <- x
rm(x)
load("./data/processed/20231112ts_ansWCp3.rda")
ts_ansWCp3 <- x
rm(x)
load("./data/processed/20231114ts_ansSCp3.rda")
ts_ansSCp3 <- x
rm(x)
load("./data/processed/20231115ts_ansSTp3.rda")
ts_ansSTp3 <- x
rm(x)
```


```{r}
ans3 <- ansb1
estim_MS <- ans3$theta
SE_MS <- sqrt(diag(ans3$covtheta))
tstat_MS <- ans3$tconv
tconv.max_MS <- ans3$tconv.max

estim_S <- ts_ansSp3$estim
SE_S <- sqrt(diag(ts_ansSp3$covtheta))
tstat_S <- ts_ansSp3$tstat
tconv.max_S <- ts_ansSp3$tconv.max

estim_WC <- ts_ansWCp3$estim
SE_WC <- sqrt(diag(ts_ansWCp3$covtheta))
tstat_WC <- ts_ansWCp3$tstat
tconv.max_WC <- ts_ansWCp3$tconv.max

estim_SC <- ts_ansSCp3$estim
SE_SC <- sqrt(diag(ts_ansSCp3$covtheta))
tstat_SC <- ts_ansSCp3$tstat
tconv.max_SC <- ts_ansSCp3$tconv.max

estim_ST <- ts_ansSTp3$estim
SE_ST <- sqrt(diag(ts_ansSTp3$covtheta))
tstat_ST <- ts_ansSTp3$tstat
tconv.max_ST <- ts_ansSTp3$tconv.max


df <- data.frame(estim_MS = estim_MS, SE_MS = SE_MS, tstat_MS = tstat_MS,
                estim_S = estim_S, SE_S = SE_S, tstat_S = tstat_S,
                 estim_WC = estim_WC, SE_WC = SE_WC, tstat_WC = tstat_WC,
                 estim_SC = estim_SC, SE_SC = SE_SC, tstat_SC = tstat_SC,
                 estim_ST = estim_ST, SE_ST = SE_ST, tstat_ST = tstat_ST)

results <- knitr::kable(df, digits = 3, "html",
             col.names = rep(c("Estim", "SE", "tstat"), 5)) %>% 
        kableExtra::add_header_above(c(" ", "ministep$^a$" = 3, "simultaneity$^b$" = 3, "weak coordination$^c$" = 3, "strict coordination$^d$" = 3, "simstep$^e$" = 3)) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
  kableExtra::add_footnote(c(paste("tconv.max:", round(tconv.max_MS, 3), sep=" "),
                             paste("tconv.max:", round(tconv.max_S, 3), sep=" "),
                             paste("tconv.max:", round(tconv.max_WC, 3), sep=" "),
                             paste("tconv.max:", round(tconv.max_SC, 3), sep=" "),
                             paste("tconv.max:", round(tconv.max_ST, 3), sep=" ")
                             ), notation="alphabet") %>%
        kableExtra::scroll_box(width = "100%", height = "500px") 
        
results
#kableExtra::save_kable(results, "./data/processed/results.html")
```
<br> 

## 5.6. GOF {.tabset .tabset-fade} 

Overal conclusion:  

- for this data and this model specification (i.e. included statistics)  
  - ministep and simultaneity are more or less identical  
  - differences between the different theories of interaction are small  
  - simstep seems to be the best overal performance  

### 5.6.1 GOF - degree distribution

```{r}
df_tsMS <- ts_degreecount(sims=ans1c_phase3$simnets, simtype="ministep") 
df_tsS <- ts_degreecount(sims=ts_ansSp3$simnets, simtype="simultaneity") 
df_tsWC <- ts_degreecount(sims=ts_ansWCp3$simnets, simtype="weak coordination") 
df_tsSC <- ts_degreecount(sims=ts_ansSCp3$simnets, simtype="strict coordination") 
df_tsST <- ts_degreecount(sims=ts_ansSTp3$simnets, simtype="simstep") 

#targets
df_target <- ts_degreecount(list(s502), simtype="target")
df_target <- df_target[,c("x", "y")]
names(df_target)[2] <- "target"
```


```{r, message=FALSE}
df <- rbind(df_tsMS, df_tsS, df_tsWC, df_tsSC, df_tsST)
df <- left_join(df, df_target)
#focus in degree until 6
df_sel <- df[df$x=="deg1" | df$x=="deg2" | df$x=="deg3"| df$x=="deg4"| df$x=="deg5"| df$x=="deg6",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary( aes(x=x, y=target, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))), fun = mean,
               geom = "point",
               color="black", shape=10, position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) + 
  labs(x = "degree", y = "degree count", fill="simulation type") 
  

p
```
<br> 
Interesting!! 

Some tentative conclusions:  

- no difference between simultaneity and ministep  
- weak coordination is best for degree 1 and degree 4. At higher degrees (5 and 6) is strict coordination the best. 
- simstep is either just as good as ministep or outperforms ministep  

### 5.6.2. GOF - dyad census  


```{r}
df_tsMS <- ts_dyads(sims=ans1c_phase3$simnets, simtype="ministep") 
df_tsS <- ts_dyads(sims=ts_ansSp3$simnets, simtype="simultaneity") 
df_tsWC <- ts_dyads(sims=ts_ansWCp3$simnets, simtype="weak coordination") 
df_tsSC <- ts_dyads(sims=ts_ansSCp3$simnets, simtype="strict coordination") 
df_tsST <- ts_dyads(sims=ts_ansSTp3$simnets, simtype="simstep") 

#targets
df_target <- ts_dyads(list(s502), simtype="target")
df_target <- df_target[,c("x", "y")]
names(df_target)[2] <- "target"
```

```{r, message=FALSE, warning=FALSE}
df <- rbind(df_tsMS, df_tsS, df_tsWC, df_tsSC, df_tsST)
df <- left_join(df, df_target)

df_sel <- df[df$x=="asym" | df$x=="mut" ,]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) + 
  stat_summary( aes(x=x, y=target, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))), fun = mean,
               geom = "point",
               color="black", shape=10, position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```
<br> 

Conclusion:  

- strict coordination seems to be the winner.  
- simstep is outperforming ministep.  

### 5.6.3. GOF - triad census  

```{r}
df_tsMS <- ts_triads(sims=ans1c_phase3$simnets, simtype="ministep") 
df_tsS <- ts_triads(sims=ts_ansSp3$simnets, simtype="simultaneity") 
df_tsWC <- ts_triads(sims=ts_ansWCp3$simnets, simtype="weak coordination") 
df_tsSC <- ts_triads(sims=ts_ansSCp3$simnets, simtype="strict coordination") 
df_tsST <- ts_triads(sims=ts_ansSTp3$simnets, simtype="simstep") 

#targets
df_target <- ts_triads(list(s502), simtype="target")
df_target <- df_target[,c("x", "y")]
names(df_target)[2] <- "target"

```

```{r}
df <- rbind(df_tsMS, df_tsS, df_tsWC, df_tsSC, df_tsST)
df <- left_join(df, df_target)

df_sel <- df[df$x=="021D" | df$x=="021U" | df$x=="021C",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) + 
  stat_summary( aes(x=x, y=target, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))), fun = mean,
               geom = "point",
               color="black", shape=10, position=position_dodge(.8)) +
  labs(x = "triad type", y = "triad count", fill="simulation type")

p

```
<br> 
Conclusion:  

- strict coordination seems to be the winner.  
- simstep is not outperforming ministep for triad configurations.  

### 5.6.4. GOF - smoke segregation  


```{r}
df_tsMS <- ts_nacf(sims=ans1c_phase3$simnets, simtype="ministep", cov=DF$smoke) 
df_tsS <- ts_nacf(sims=ts_ansSp3$simnets, simtype="simultaneity", cov=DF$smoke) 
df_tsWC <- ts_nacf(sims=ts_ansWCp3$simnets, simtype="weak coordination", cov=DF$smoke) 
df_tsSC <- ts_nacf(sims=ts_ansSCp3$simnets, simtype="strict coordination", cov=DF$smoke) 
df_tsST <- ts_nacf(sims=ts_ansSTp3$simnets, simtype="simstep", cov=DF$smoke) 

#targets
df_target <- ts_nacf(list(s502), simtype="target", cov=DF$smoke)
names(df_target)[1] <- "target"

```

```{r}
df <- rbind(df_tsMS, df_tsS, df_tsWC, df_tsSC, df_tsST)
names(df)[1] <- "MoranI" 
df$target <- df_target$target

p <- ggplot(df, aes(x=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep")), y=MoranI, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) + 
  stat_summary( aes(x=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep")), y=target, fill=factor(type, levels=c("ministep", "simultaneity", "weak coordination", "strict coordination", "simstep"))), fun = mean,
               geom = "point",
               color="black", shape=10, position=position_dodge(.8)) +
  labs(x = "simulation type", y = "smoke segregation (MoranI)", fill="simulation type") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

p

```
<br> 
Conclusion:  

- weak and strict coordination are doing worst  

# 6. Conclusion  

RsienaTwoStep offers a workflow for assessing the extent to which the ministep assumption is crucial. By crucial I mean that parameter estimates and model fit depend on the chosen 'micro theory of interaction'.  
In the above example, the assumption is not crucial. All theories of interaction lead to similar conclusions with respect to the significance of the included statistics and the GOF of the ministep and twostep models are very similar. Perhaps the simstep models has the best GOF.  




---- 

- klad



# okay let us re-estimate the simulated data

```{r}
ts_simdata$startvalues_sim
friendship <- sienaDependent(array(c(ts_simdata$net_sim1, ts_simdata$net_sim2), dim = c(50, 50,
    2)))  

alcohol <- sienaDependent(as.matrix(ts_simdata$DF_sim[,1:2]), type = "behavior")
smoke <- coCovar(as.numeric(ts_simdata$DF_sim[,3]))
mydata <- sienaDataCreate(friendship, smoke, alcohol)

myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, cycle3, transTrip)
myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
myeff <- includeEffects(myeff, absDiffX, interaction1 = "alcohol")
myeff <- includeEffects(myeff, name = "alcohol", avAlt, interaction1 = "friendship")
myeff <- includeEffects(myeff, name = "alcohol", effFrom, interaction1 = "smoke")

print01Report(mydata)

myalgorithm <- sienaAlgorithmCreate(cond = FALSE, findiff = TRUE, projname="simdata") 

anssimdata1 <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, returnDeps = TRUE)
ans2simdata1 <- siena07(myalgorithm, data=mydata, effects=myeff, prevAns = anssimdata1, batch=TRUE, returnDeps = TRUE)
```

```{r}

```



```{r}
library(igraph)
g1 <- (graph_from_adjacency_matrix(ts_simdata$net_sim1, mode = "directed"))
V(g1)$size <- .2
E(g1)$arrow.size <- .2
l <- layout_with_mds(g1)
plot(g1, layout = l)


g2 <- (graph_from_adjacency_matrix(ts_simdata$net_sim2, mode = "directed"))
V(g2)$size <- .2
E(g2)$arrow.size <- .2
plot(g2, layout = l)
```


