---
title: "RSienaWorkFlow II"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.test {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
}

.test2 {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 10px;
  background-color: white;
  color: rgb(201, 76, 76);
}


h1, .h1, h2, .h2, h3, .h3 {
  margin-top: 24px;
}





.button1 {
  background-color: grey; /* Red */ 
  border: 2px solid black;
  color: white;
  padding: 15px 32px;
  text-align: center;
  text-decoration: none;
  display: inline-block;
  font-size: 16px;
  margin: 4px 2px;
  cursor: pointer;
  /* border-radius: 12px; */
  width: 100%;
}

.button1:hover {
  box-shadow: 0 12px 16px 0 rgba(0,0,0,0.24), 0 17px 50px 0 rgba(0,0,0,0.19);
}

.button1:active {
  border: 2px solid red;
}

pre {
  max-height: 300px;
  overflow-y: auto;
  overflow-x: auto;
  margin: 0px;
}

.watch-out {
  background-color: lightpink;
  border: 3px solid red;
  font-weight: bold;
}

.large {
  max-height: 1100px;
  overflow-y: auto;
  overflow-x: auto;
  width: 1100px;
}


body{ /* Normal  */
    font-size: 14px;
}

h1 { /* Header 1 */
    font-size: 20px;
  font-weight: bold;
}

h2 { /* Header 2 */
    font-size: 18px;
  color: DarkBlue;
}

h3 { /* Header 3 */
    font-size: 16px;
  color: DarkBlue;
}

blockquote {
  padding: 10px 20px;
  margin: 0 0 20px;
  font-size: 14px;
  border-left: 6px solid #eee;
  background-color:lightcyan
}

caption {
  font-size: 14px;
  color: black;
  font-weight: bold
}

pre.numberSource a.sourceLine {
  left: -1em;
}

```


# 1. Getting started


```{r}
rm(list=ls())
```


## 1.1. Basic functions  


```{r}
fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}

```

## 1.2. Packages

```{r}
packages = c("RsienaTwoStep", "RSiena", "doParallel", "compiler", "ggplot2", "metafor")

fpackage.check(packages)
```

## 1.3. Auxiliary functions for `RSiena::SienaGOF()`

Please see: `RSiena::SienaGOF()`.  

I want to be able to have a dyadcensus/triadcensus in the GOF of Rsiena. 
Thus, need to add this. 

```{r}

DyadCensus.sna <- function(i, data, sims, period, groupName, varName, levls=1:3){
  unloadNamespace("igraph") # to avoid package clashes
  require(network)
  require(sna)
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}
  # because else triad.census(x) will lead to an error
  tc <- sna::dyad.census(x)[levls]
  # names are transferred automatically
  tc
}

DyadCensus.sna <- cmpfun(DyadCensus.sna)


TriadCensus.sna <- function(i, data, sims, period, groupName, varName, levls=1:16){
  unloadNamespace("igraph") # to avoid package clashes
  require(network)
  require(sna)
  x <- networkExtraction(i, data, sims, period, groupName, varName)
  if (network.edgecount(x) <= 0){x <- symmetrize(x)}
  # because else triad.census(x) will lead to an error
  tc <- sna::triad.census(x)[levls]
  # names are transferred automatically
  tc
}

TriadCensus.sna <- cmpfun(TriadCensus.sna)

```

---  

# 2. Setting up cluster

```{r}
#stopCluster(my.cluster) 
n.cores <- parallel::detectCores() - 2  #save one core for other work
# create the cluster
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK")
# register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)
```

---  


# 4. Model 2: Degree, reciprocity and 3cycle 

## 4.1. Running `siena07()`  

We only include the following statistics:  
- degree  
- reciprocity  
- cycle3  


```{r, eval=FALSE}
mynet <- sienaDependent(array(c(s501, s502), dim=c(50, 50, 2)))
mydata <- sienaDataCreate(mynet)
myalgorithm <- sienaAlgorithmCreate(seed=1293, cond=FALSE) #toggle set conditional to retrieve the rate parameter in theta!
myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, cycle3)
ans2 <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, returnDeps = TRUE)
```


```{r, eval=FALSE}
fsave(ans2, file="ans2.Rdata")
```

```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221222ans2.Rdata")
ans2 <- x
rm(x)
```

let's have a look
```{r}
ans2
```


## 4.2. Simulate networks via `RsienaTwoStep`  

`r colorize("These simulations took two days to complete on a simple HP Elitebook 855 G7", "red" )`

```{r, eval=FALSE, class.source="watch-out"}
#ministep
sims1m2 <- ts_sims(nsims=1000, parallel=TRUE, net=s501, rate=summary(ans2)$theta[1], statistics=list(ts_degree, ts_recip, ts_cycle3), parameters=summary(ans2)$theta[2:4], p2step=c(1,0,0), chain=FALSE)
fsave(sims1m2, file="sims1m2.Rdata")

#simultaneity
sims2m2 <- ts_sims(nsims=1000, parallel=TRUE, net=s501, rate=summary(ans2)$theta[1], statistics=list(ts_degree, ts_recip, ts_cycle3), parameters=summary(ans2)$theta[2:4], p2step=c(0,1,0), chain=FALSE)
fsave(sims2m2, file="sims2m2.Rdata")

#weak coordination
sims3m2 <- ts_sims(nsims=1000, parallel=TRUE, net=s501, rate=summary(ans2)$theta[1], statistics=list(ts_degree, ts_recip, ts_cycle3), parameters=summary(ans2)$theta[2:4], p2step=c(0,1,0), dist1=2, dist2=2, chain=FALSE)
fsave(sims3m2, file="sims3m2.Rdata")

#strict coordination
sims4m2 <- ts_sims(nsims=1000, parallel=TRUE, net=s501, rate=summary(ans2)$theta[1], statistics=list(ts_degree, ts_recip, ts_cycle3), parameters=summary(ans2)$theta[2:4], p2step=c(0,1,0), dist1=2, chain=FALSE)
fsave(sims4m2, file="sims4m2.Rdata")

#simstep
sims5m2 <- ts_sims(nsims=1000, parallel=TRUE, net=s501, rate=summary(ans2)$theta[1], statistics=list(ts_degree, ts_recip, ts_cycle3), parameters=summary(ans2)$theta[2:4], p2step=c(0,0,1), chain=FALSE)
fsave(sims5m2, file="sims5m2.Rdata")

```


```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20221222sims1m2.Rdata")
sims1m2 <- x
load("./data/processed/20221223sims2m2.Rdata")
sims2m2 <- x
load("./data/processed/20221223sims3m2.Rdata")
sims3m2 <- x
load("./data/processed/20221224sims4m2.Rdata")
sims4m2 <- x
load("./data/processed/20231225sims5m2.Rdata")
sims5m2 <- x
rm(x)
```

## 4.3. Counting dyads

### 4.3.1. RSiena

```{r}
gofi <- sienaGOF(ans2,
                 DyadCensus.sna, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "mynet")
```

```{r}
dfrs <- as.data.frame(gofi[[1]]$Simulations)
names(dfrs) <- c("Mut", "Asym", "Null")

df <- rbind(dfrs, dfrs, dfrs)
df$x <- rep(c("mut", "asym", "null"), each=nrow(dfrs))
df$y <- NA
df$y[df$x=="mut"] <- df$Mut[df$x=="mut"]
df$y[df$x=="asym"] <- df$Asym[df$x=="asym"]
df$y[df$x=="null"] <- df$Null[df$x=="null"]
df$type <- "rsiena"
dfrs_sienam2 <- df
rm(list=c("df", "dfrs"))
```

### 4.3.2. RsienaTwoStep 

```{r}
df_ts1m2 <- ts_dyads(sims=sims1m2, simtype="ministep") 
df_ts2m2 <- ts_dyads(sims=sims2m2, simtype="twostep-simultaneity")
df_ts3m2 <- ts_dyads(sims=sims3m2, simtype="twostep-weak coordination") 
df_ts4m2 <- ts_dyads(sims=sims4m2, simtype="twostep-strict coordination")
df_ts5m2 <- ts_dyads(sims=sims5m2, simtype="simultaneity")
```

## 4.4. Plotting dyad_census

### 4.4.1. All three dyad types  

```{r}
df <- rbind(dfrs_sienam2, df_ts1m2, df_ts2m2, df_ts3m2, df_ts4m2, df_ts5m2)

p <- ggplot(df, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simultaneity"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```

### 4.4.2. Plotting dyad census selection 
```{r}
df_sel <- df[df$x!="null", ]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simultaneity"))) ) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "dyad type", y = "dyad count", fill="simulation type")

p

```


## 4.5. Counting triads

### 4.5.1. RSiena
```{r}
nsims <- 1000
gofi <- sienaGOF(ans2,
                 TriadCensus.sna, 
                 verbose = TRUE,
                 join = TRUE, 
                 varName = "mynet")

dfrs <- as.data.frame(gofi[[1]]$Simulations)

triads <- c("003",  "012",  "102" , "021D", "021U", "021C" ,"111D", "111U" ,"030T" ,"030C", "201"  ,"120D" ,"120U" ,"120C" ,"210",  "300")

names(dfrs) <- triads


dflist <- list()
for (i in 1:length(triads)) {
  dflist[[i]] <- dfrs
}
dfrs <- do.call(rbind, dflist)
dfrs$x <- rep(triads, each=nsims)
dfrs$y <- NA

for (i in 1:length(triads)) {
  dfrs$y[dfrs$x==triads[i]] <- dfrs[,triads[i]][dfrs$x==triads[i]]
}
dfrs$type <- "rsiena"
dfrs_sienam2 <- dfrs
```


### 4.5.2 RsienaTwoStep

```{r}
df_ts1m2 <- ts_triads(sims=sims1m2, simtype="ministep") 
df_ts2m2 <- ts_triads(sims=sims2m2, simtype="twostep-simultaneity")
df_ts3m2 <- ts_triads(sims=sims3m2, simtype="twostep-weak coordination") 
df_ts4m2 <- ts_triads(sims=sims4m2, simtype="twostep-strict coordination")
df_ts5m2 <- ts_triads(sims=sims5m2, simtype="simstep")
```


## 4.6. Plotting triad census 

### 4.6.1. All triad types  

```{r}
df <- rbind(dfrs_sienam2, df_ts1m2, df_ts2m2, df_ts3m2, df_ts4m2, df_ts5m2)

p <- ggplot(df, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simultaneity")))) + 
  geom_violin(position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(1)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(1)) +
  labs(x = "triad type", y = "triad count", fill="simulation type")

p
```

### 4.6.2 Plotting triad census selection I

```{r}
df_sel <- df[df$x=="030C" | df$x=="120C",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simultaneity")))) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "(cyclic) triad types", y = "triad count", fill="simulation type")

p
```

### 4.6.3 Plotting triad census selection II

```{r}
df_sel <- df[df$x=="210" | df$x=="300",]

p <- ggplot(df_sel, aes(x=x, y=y, fill=factor(type, levels=c("rsiena", "ministep", "twostep-simultaneity", "twostep-weak coordination", "twostep-strict coordination", "simultaneity")))) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "(cyclic) triad types", y = "triad count", fill="simulation type")

p
```

---  


# Re-estimation 

# let us estimate the simulated model of twostep-weak coordination again via Rsiena

```{r}
mynet <- sienaDependent(array(c(s501, sims3m2[[1]]), dim=c(50, 50, 2)))
mydata <- sienaDataCreate(mynet)
myalgorithm <- sienaAlgorithmCreate(seed=1293, cond=FALSE) #toggle set conditional to retrieve the rate parameter in theta!
myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, cycle3)


tests1 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims1m2[[i]]), dim=c(50, 50, 2)))
  mydata <- sienaDataCreate(mynet)
  ans <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se)
}

tests2 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims2m2[[i]]), dim=c(50, 50, 2)))
  mydata <- sienaDataCreate(mynet)
  ans <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se)
}

tests3 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims3m2[[i]]), dim=c(50, 50, 2)))
  mydata <- sienaDataCreate(mynet)
  ans <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se)
}

tests4 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims4m2[[i]]), dim=c(50, 50, 2)))
  mydata <- sienaDataCreate(mynet)
  ans <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se)
}

tests5 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims5m2[[i]]), dim=c(50, 50, 2)))
  mydata <- sienaDataCreate(mynet)
  ans <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se)
}


```

pooled effects

```{r}
library(metafor)


test <- tests1
rate <- rma(yi=test[,1], sei=test[,5], method="FE")
degree <- rma(yi=test[,2], sei=test[,6], method="FE")
reciprocity <- rma(yi=test[,3], sei=test[,7], method="FE")
cycle3 <- rma(yi=test[,4], sei=test[,8], method="FE")

rate$se
estimates <- rbind(rate$b, degree$b, reciprocity$b, cycle3$b)
se <- rbind(rate$se, degree$se, reciprocity$se, cycle3$se)

df <- data.frame(Estimate=estimates, "Standard Error"=se)
rownames(df) <- c("rate", "outdegree", "reciprocity", "3-cycles")
dfs1 <- df

test <- tests2
rate <- rma(yi=test[,1], sei=test[,5], method="FE")
degree <- rma(yi=test[,2], sei=test[,6], method="FE")
reciprocity <- rma(yi=test[,3], sei=test[,7], method="FE")
cycle3 <- rma(yi=test[,4], sei=test[,8], method="FE")

rate$se
estimates <- rbind(rate$b, degree$b, reciprocity$b, cycle3$b)
se <- rbind(rate$se, degree$se, reciprocity$se, cycle3$se)

df <- data.frame(Estimate=estimates, "Standard Error"=se)
rownames(df) <- c("rate", "outdegree", "reciprocity", "3-cycles")
dfs2 <- df

test <- tests3
rate <- rma(yi=test[,1], sei=test[,5], method="FE")
degree <- rma(yi=test[,2], sei=test[,6], method="FE")
reciprocity <- rma(yi=test[,3], sei=test[,7], method="FE")
cycle3 <- rma(yi=test[,4], sei=test[,8], method="FE")

rate$se
estimates <- rbind(rate$b, degree$b, reciprocity$b, cycle3$b)
se <- rbind(rate$se, degree$se, reciprocity$se, cycle3$se)

df <- data.frame(Estimate=estimates, "Standard Error"=se)
rownames(df) <- c("rate", "outdegree", "reciprocity", "3-cycles")
dfs3 <- df

test <- tests4
rate <- rma(yi=test[,1], sei=test[,5], method="FE")
degree <- rma(yi=test[,2], sei=test[,6], method="FE")
reciprocity <- rma(yi=test[,3], sei=test[,7], method="FE")
cycle3 <- rma(yi=test[,4], sei=test[,8], method="FE")

rate$se
estimates <- rbind(rate$b, degree$b, reciprocity$b, cycle3$b)
se <- rbind(rate$se, degree$se, reciprocity$se, cycle3$se)

df <- data.frame(Estimate=estimates, "Standard Error"=se)
rownames(df) <- c("rate", "outdegree", "reciprocity", "3-cycles")
dfs4 <- df

test <- tests5
rate <- rma(yi=test[,1], sei=test[,5], method="FE")
degree <- rma(yi=test[,2], sei=test[,6], method="FE")
reciprocity <- rma(yi=test[,3], sei=test[,7], method="FE")
cycle3 <- rma(yi=test[,4], sei=test[,8], method="FE")

rate$se
estimates <- rbind(rate$b, degree$b, reciprocity$b, cycle3$b)
se <- rbind(rate$se, degree$se, reciprocity$se, cycle3$se)

df <- data.frame(Estimate=estimates, "Standard Error"=se)
rownames(df) <- c("rate", "outdegree", "reciprocity", "3-cycles")
dfs5 <- df

```


```{r}
ans2
round(dfs1,4)
round(dfs2,4)
round(dfs3,4)
round(dfs4,4)
round(dfs5,4)
```


---  

# Model with segregtion 

## setting up the cluster
```{r}
#stopCluster(my.cluster) 
n.cores <- parallel::detectCores() - 2  #save one core for other work
# create the cluster
my.cluster <- parallel::makeCluster(n.cores, type = "PSOCK")
# register it to be used by %dopar%
doParallel::registerDoParallel(cl = my.cluster)
```


## load the simulated networks


```{r, eval=TRUE, echo=FALSE}
load("./data/processed/20230501sims1m3.Rdata")
sims1m3 <- x
load("./data/processed/20230507sims1m3b.Rdata")
sims1m3b <- x
load("./data/processed/20230507sims2m3.Rdata")
sims2m3 <- x
load("./data/processed/20230507sims3m3.Rdata")
sims3m3 <- x
load("./data/processed/20230507sims4m3.Rdata")
sims4m3 <- x
load("./data/processed/20230507sims5m3.Rdata")
sims5m3 <- x
rm(x)
```


## setting up algorithms

We do this both for the normal MiniStep and the ReactiveStep

```{r}
myalgorithm1 <- sienaAlgorithmCreate(cond=FALSE)
myalgorithm2 <- sienaAlgorithmCreate(cond=FALSE, modelType=c(mynet=10))
```

### Actual re-estimation 
```{r}
REsims1m3b <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims1m3b[[i]]), dim=c(50, 50, 2)))
  alcohol <- s50a
  smoke <- s50s
  smoke <- coCovar(smoke[, 1])
  alcohol <- coCovar(alcohol[, 1])
  mydata <- sienaDataCreate(mynet, smoke, alcohol)

  myeff <- getEffects(mydata)
  myeff <- includeEffects(myeff, transTrip, cycle3)
  myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
  myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
  
  ans <- siena07(myalgorithm1, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  ans2 <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se, ans2$theta, ans2$se)
}

fsave(REsims1m3b, file="REsims1m3b.Rdata")

```


```{r}
REsims2m3 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims2m3[[i]]), dim=c(50, 50, 2)))
  alcohol <- s50a
  smoke <- s50s
  smoke <- coCovar(smoke[, 1])
  alcohol <- coCovar(alcohol[, 1])
  mydata <- sienaDataCreate(mynet, smoke, alcohol)

  myeff <- getEffects(mydata)
  myeff <- includeEffects(myeff, transTrip, cycle3)
  myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
  myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
  
  ans <- siena07(myalgorithm1, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  ans2 <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se, ans2$theta, ans2$se)
}

fsave(REsims2m3, file="REsims2m3.Rdata")


REsims3m3 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims3m3[[i]]), dim=c(50, 50, 2)))
  alcohol <- s50a
  smoke <- s50s
  smoke <- coCovar(smoke[, 1])
  alcohol <- coCovar(alcohol[, 1])
  mydata <- sienaDataCreate(mynet, smoke, alcohol)

  myeff <- getEffects(mydata)
  myeff <- includeEffects(myeff, transTrip, cycle3)
  myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
  myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
  
  ans <- siena07(myalgorithm1, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  ans2 <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se, ans2$theta, ans2$se)
}

fsave(REsims3m3, file="REsims3m3.Rdata")

REsims4m3 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims4m3[[i]]), dim=c(50, 50, 2)))
  alcohol <- s50a
  smoke <- s50s
  smoke <- coCovar(smoke[, 1])
  alcohol <- coCovar(alcohol[, 1])
  mydata <- sienaDataCreate(mynet, smoke, alcohol)

  myeff <- getEffects(mydata)
  myeff <- includeEffects(myeff, transTrip, cycle3)
  myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
  myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
  
  ans <- siena07(myalgorithm1, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  ans2 <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se, ans2$theta, ans2$se)
}
fsave(REsims4m3, file="REsims4m3.Rdata")

REsims5m3 <- foreach(i=1:100, .combine=rbind, .packages="RSiena") %dopar% {
  mynet <- sienaDependent(array(c(s501, sims5m3[[i]]), dim=c(50, 50, 2)))
  alcohol <- s50a
  smoke <- s50s
  smoke <- coCovar(smoke[, 1])
  alcohol <- coCovar(alcohol[, 1])
  mydata <- sienaDataCreate(mynet, smoke, alcohol)

  myeff <- getEffects(mydata)
  myeff <- includeEffects(myeff, transTrip, cycle3)
  myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
  myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
  
  ans <- siena07(myalgorithm1, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  ans2 <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, silent=TRUE)
  c(ans$theta, ans$se, ans2$theta, ans2$se)
}
fsave(REsims5m3, file="REsims5m3.Rdata")
```


```{r}
load("./data/processed/20230508REsims1m3b.Rdata")
REsims1m3b <- x
load("./data/processed/20230508REsims2m3.Rdata")
REsims2m3 <- x
load("./data/processed/20230508REsims3m3.Rdata")
REsims3m3 <- x
load("./data/processed/20230508REsims4m3.Rdata")
REsims4m3 <- x
load("./data/processed/20230508REsims5m3.Rdata")
REsims5m3 <- x

```



## pool the effects

```{r}

REsims <- list(REsims1m3b, REsims2m3, REsims3m3, REsims4m3, REsims5m3)
df_list <- list()

for (i in 1:5) {
  
  test <- REsims[[i]]
  
  #ministep
  rate1 <- rma(yi=test[,1], sei=test[,10], method="FE")
  degree1 <- rma(yi=test[,2], sei=test[,11], method="FE")
  reciprocity1 <- rma(yi=test[,3], sei=test[,12], method="FE")
  transitive_triplets1 <- rma(yi=test[,4], sei=test[,13], method="FE")
  cycle31 <- rma(yi=test[,5], sei=test[,14], method="FE")
  smoke_similarity1 <- rma(yi=test[,6], sei=test[,15], method="FE")
  alcohol_alter1 <- rma(yi=test[,7], sei=test[,16], method="FE")
  alcohol_ego1 <- rma(yi=test[,8], sei=test[,17], method="FE")
  alcohol_egoXalcohol_alter1 <- rma(yi=test[,9], sei=test[,18], method="FE")
  
  #reactivestep
  rate2 <- rma(yi=test[,19], sei=test[,28], method="FE")
  degree2 <- rma(yi=test[,20], sei=test[,29], method="FE")
  reciprocity2 <- rma(yi=test[,21], sei=test[,30], method="FE")
  transitive_triplets2 <- rma(yi=test[,22], sei=test[,31], method="FE")
  cycle32 <- rma(yi=test[,23], sei=test[,32], method="FE")
  smoke_similarity2 <- rma(yi=test[,24], sei=test[,33], method="FE")
  alcohol_alter2 <- rma(yi=test[,25], sei=test[,34], method="FE")
  alcohol_ego2 <- rma(yi=test[,26], sei=test[,35], method="FE")
  alcohol_egoXalcohol_alter2 <- rma(yi=test[,27], sei=test[,36], method="FE")
  
  
  
  estimates1 <- rbind(rate1$b, degree1$b, reciprocity1$b, transitive_triplets1$b, cycle31$b, smoke_similarity1$b,alcohol_alter1$b, alcohol_ego1$b, alcohol_egoXalcohol_alter1$b)
  
  se1 <- rbind(rate1$se, degree1$se, reciprocity1$se, transitive_triplets1$se, cycle31$se, smoke_similarity1$se, alcohol_alter1$se, alcohol_ego1$se, alcohol_egoXalcohol_alter1$se)
  
  estimates2 <- rbind(rate2$b, degree2$b, reciprocity2$b, transitive_triplets2$b, cycle32$b, smoke_similarity2$b,alcohol_alter2$b, alcohol_ego2$b, alcohol_egoXalcohol_alter2$b)
  
  se2 <- rbind(rate2$se, degree2$se, reciprocity2$se, transitive_triplets1$se, cycle32$se, smoke_similarity2$se, alcohol_alter2$se, alcohol_ego2$se, alcohol_egoXalcohol_alter2$se)
  
  
  
  df <- data.frame("Estimate (MiniStep)"=estimates1, "Standard Error"=se1, "Estimate (ReactiveStep)"=estimates2, "Standard Error"=se2)
  rownames(df) <- c("rate", "degree", "reciprocity", "transitive_triplets", "cycle3", "smoke_similarity", "alcohol_alter", "alcohol_ego", "alcohol_egoXalcohol_alter")
  
  df_list[[i]] <- df
  
}
round(df_list[[5]],2)
```



# comparing Ministep with ReactiveStep

## 5.1. Running `Siena07()

preparing the dataset
```{r}
mynet <- sienaDependent(array(c(s501, s502), dim=c(50, 50, 2)))
alcohol <- s50a
smoke <- s50s
smoke <- coCovar(smoke[, 1])
alcohol <- coCovar(alcohol[, 1])
mydata <- sienaDataCreate(mynet, smoke, alcohol)

myeff <- getEffects(mydata)
myeff <- includeEffects(myeff, transTrip, cycle3)
myeff <- includeEffects(myeff, egoX, altX, egoXaltX, interaction1 = "alcohol")
myeff <- includeEffects(myeff, simX, interaction1 = "smoke")
```

### ministep
```{r, eval=FALSE}
myalgorithm <- sienaAlgorithmCreate(seed=1293, cond=FALSE) #toggle set conditional to retrieve the rate parameter in theta!

ans3 <- siena07(myalgorithm, data=mydata, effects=myeff, batch=TRUE, returnDeps = TRUE)
ans3
```

```{r}
load("./data/processed/20230401ans3.Rdata")
ans3 <- x
rm(x)
ans3
```



### ReactiveStep

```{r}
?sienaAlgorithm
myalgorithm2 <- sienaAlgorithmCreate(seed=1293, cond=FALSE, modelType=c(mynet=10)) 
                                     
                                     
                                     #toggle set conditional to retrieve the rate parameter in theta!
ans3RS <- siena07(myalgorithm2, data=mydata, effects=myeff, batch=TRUE, returnDeps = TRUE)
ans3RS                                     
                                     
```

```{r}
fsave(ans3RS, "ans3RS")
```

```{r}
load("./data/processed/20230507ans3RS")
ans3RS <- x
rm(x)
ans3RS
```

# test for AF


```{r}
net1 <- matrix(0, 20, 20)
group <- rep(c(1:2), 10)
```

```{r}
sims0 <- ts_sims(nsims=20, parallel=TRUE, net=net1, ccovar = data.frame(group=group), rate=100, statistics=list(ts_degree, list(ts_sameX, "group")), parameters=c(-1,1), p2step=c(1,0,0), chain=FALSE) #ministep
sims1 <- ts_sims(nsims=20, parallel=TRUE, net=net1, ccovar = data.frame(group=group), rate=100, statistics=list(ts_degree, list(ts_sameX, "group")), parameters=c(-1,1), p2step=c(0,1,0), chain=FALSE) #twostep simultaneity
sims2 <- ts_sims(nsims=1, parallel=FALSE, net=net1, ccovar = data.frame(group=group), rate=100, statistics=list(ts_degree, list(ts_sameX, "group")), parameters=c(-1,1), p2step=c(0.1,.9,0), dist1=1, chain=FALSE) #strict coordination ##why is this one not working!
```



```{r}
seg0 <- ts_nacf(sims=sims1, simtype="ministep", cov=group) 
names(seg0)[1] <- "MoranI"
seg1 <- ts_nacf(sims=sims1, simtype="simultaneity", cov=group) 
names(seg1)[1] <- "MoranI" 
seg2 <- ts_nacf(sims=sims2, simtype="simultaneity/twostep", cov=group) 
names(seg2)[1] <- "MoranI" 
```



```{r}
df <- rbind(seg0, seg1, seg2)


p <- ggplot(df, aes(x=factor(type , levels=c("ministep", "simultaneity", "simultaneity/twostep" )), y=MoranI, fill=factor(type, levels=c("ministep", "simultaneity", "simultaneity/twostep")))) + 
  geom_violin(position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "errorbar",
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               width=.1,
               color="red", position=position_dodge(.8)) + 
  stat_summary(fun = mean,
               geom = "point",
               color="red", position=position_dodge(.8)) +
  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type")  + 
  ylim(-1, 1) + 
  theme(axis.text.x=element_blank())
        
#  labs(x = "sim type", y = "Smoke \n MoranI", fill="sim type") 
#+
#  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

p

```

